---
title: "Xaringan project"
author: "Alejandro Molina-Moctezuma"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download the whole project  

Download it from: https://github.com/moctezumaii/Xar_project    

You can download the whole Zip file(click on code, then download Zip file). Make sure that when you extract the zip file, everything is saved in the same place.   

The file (already knitted) is under: Presentation.html. Slides will auto switch every 20 seconds (you can also switch them manually)  


## Instructions to run project    

Download Devtools.    
Go to:
https://cran.r-project.org/bin/windows/Rtools/    
  
and click on download (either 64 or 32 bit), and install it.
If that doesn't work you can also try:    

```{r, eval=F, echo=T}
install.packages('devtools')
```

Install Xaringan. Best way:

```{r, eval=F, echo=T}
devtools::install_github('yihui/xaringan') 
```

Alternative:    

```{r, eval=F, echo=T}
install.packages('xaringan')
```




## Download fonts:    
You don't need to download the fonts, as I included them in the folder (they are under assets). So you can skip this step.   
Here are the instructions, just in case (again, you can skip this step)   
Download fonts from: https://github.com/garthtarr/sydney_xaringan
You can download the whole Zip file(click on code, then download Zip file), or you can download the specific files:    

sydney-fonts.css    
sydney.css    
mathjax-equation-numbers.html   
garthtarr/sydney_xaringan       



## Install other packages    
```{r, eval=F, echo=T}
remotes::install_github("mitchelloharawild/icons")
install.packages("RefManageR")
install.packages("DT")
install.packages("tidyverse")
install.packages("gridExtra")
install.packages("cowplot")
```
  
Installing the icons might be problematic. If you get an error saying: could not find package 'xxx', then you need to install that package using install.package['xxx']. This may happen for multiple packages, but it should be easy to solve past that.
  
## Necessary files    
Please, do have all of the following files in the same file as the Presentation.Rmd file. This should be the case if you downloaded everything in the same place in your computer.    

Files needed for slide 16:  

* df_comparisions.RDS   
* Overall_feature_importance.RDS    

Files used in other slides:   

* shot_logs.csv
* xg_model_2Pts.rds
* clean.csv   

## Running it
It should run now. Just click "knit", and it will run. Do note that in the bottom-right corner, there is a timer.      
After you run it, it will save an html file and the original one will be overwritten. 


## Additional help    
If you need any modifications contact me through Upwork or send an email to amolinamoctezuma@lssu.edu
